\begin{abstract}
  There are many user-level APIs for connecting and exchanging data
  between network peers.  Each interface varies wildly when evaluating
  them based on metrics including (but not limited to) performance,
  portability, and complexity.  Specifically, many interfaces make
  design or implementation choices that end up emphasizing some of the
  more desirable metrics (e.g., performance) while sacrificing others
  (e.g., portability).  As a direct result, software developers
  building large, network-based applications end up having to choose
  which network API to use based on a complex, multi-dimensional set
  of criteria -- but still inevitably end up with an interface that
  fails to deliver features in at least some desirable metrics.

  In this paper, we introduce a novel interface that both supports
  many features that have become standard (or otherwise generally
  expected) in other communication interfaces, and strives to export a
  small, yet powerful, interface.  This new interface draws upon years
  of diverse networking research spanning from theories of application
  software development to systems-level implementation algorithms.
  The goal was to create a relatively simple, high-level
  communications interface -- one that might even feel somewhat
  ``familiar'' to developers because of its roots in many other,
  existing network APIs -- that still provides important features such
  as scalability, resiliency, and performance.  The result is the
  Common Communications Interface (CCI): an API that is simple enough
  to implement in hardware, explicitly designed to incur minimal
  overhead from a host operating system, and performant enough to meet
  the needs of network-intensive applications.
\end{abstract}

% LocalWords:  performant
